Search for Spatially Extended \fermi-LAT Sources Using Two Years of Data

### 2012-04-25 extended_catalog/external_review_reply/replies_to_review_comments.tex

\begin{referee}
Major issues :

The authors use a uniform disk model for fitting the extended sources.
However many sources show a complex morphology (for example, not radial symmetric)
They should quantify how the results they are presenting depend on the actual shape of the sources.
\end{referee}

\begin{reply}
In the new draft, Section 3 (Validation of Analysis Method) has been
divided into two sections. The original section is now section 3.1
(now called Statistical Significance Testing).

There is a new section 3.2 (Assessing the Bias Due to Incorrect
Spatial Models). In this section, we address your question by performing
a dedicated Monte Carlo study of W44 simulated with a ring-type spatial
model. By fitting the source with different spatial models, we asses
this bias and find it to be very small.

@@

This new section requires two new figures (Figure 6 and Figure 7). Also,
in section 2.4 (Comparing Source Size) we add a line of text referring
to this new section. Finally, there is a new sentance in the introduction
of the paper which references the new text.
\end{reply}

\begin{referee}
Most of the new extended sources found in this work lie on the galactic
plane at very low latitudes. Why the simulations used to validate the
analysis (par. 3) do not include the diffuse Galactic background? Can
the structures of this diffuse background affect the analysis results?
\end{referee}

\begin{reply}
In section 3.1 (which was previously just section 3 and is now called
Statistical Significance Testing), we have added new text describing a
follow up Monte Carlo study performed in the presence of the galactic
plane. In addition, this new text required updating Table 1 (Monte Carlo
Spectral Parameters) to include the new simulation parameters. There is also
a new figure (Figure 5) which shows the distribution of TSext for these
new simulations.
\end{reply}

@@

\begin{referee}
An important result of this work is that some 2FGL sources could be fake due to the residuals induced by modeling the extended sources as point-like. It may be worth having a table that lists the sources that are not longer needed in the modeling of the sky (alternatively this information could be included in table 4). They must also be mentioned in the captions of figures 16, 17, 19, 21 and 22.
\end{referee}

\begin{reply}
We have added a new table (Table 6 in the current draft, named Nearby Residual-induced Sources) which contains
this information. There are paragraphs in section 9 (Extended Source
Search Method) and Section 10 (New Extended Sources) which refer to this
table. We have also updated the requested captions to list names of 2FGL
sources that were removed.
\end{reply}

@@

Model selection today is instead guided by `penalized' measures,
and the most widely used is the Bayesian Information Criterion
(BIC,\url{http://en.wikipedia.org/wiki/Bayesian_information_criterion}). It
is very easy to compute, essentially an offset from the LRT scaled to
the log of the counts involved. The issues are discussed in detail in
the volume ``Model Selection and Multi-Model Inference'' (K.P. Burnham
\& D. Anderson, 2nd ed, 2002); see also lectures (especially \#14) at
\url{http://myweb.uiowa.edu/cavaaugh/ms_seminar.html}.

I guess you could argue that your Monte Carlo experiments (Figs 3-4) take
the flux and degrees of freedom into account for the LRT statistic. But
if you want your method to be widely used, it should not rest on a weak
and obsolete statistical foundation. Altogether, I recommend moving to
the BIC. Your approach of Monte Carlo testing of the significance of
LRT values should also work for BIC values.
\end{referee}

\begin{reply}
Concerning the extension significance test, we extensively validated it
against Monte Carlo simulations. Furthermore, the high energy astrophysics
community has a lot of experience with the test statistic (TS) applied to
nested models. So we are very confident that we can apply it to the actual
data.

Concerning the test comparing two point-like sources to one extended
source, we find the desire for a Bayesian test more compelling since the
models are not nested. We considered the BIC test but found that it would
not work well for LAT data. We alternately considered the AIC test and
found that it is equivalent to the test we actually performed. We have
added a new paragraph to Section 5 (Testing Against Source Confusion)
that describes the Bayesian interpretation of the test we actually performed.
\end{reply}

### 2012-05-08 extended_catalog/external_review_reply/replies_to_review_comments.tex

\title{Reply to Referee Report}

\begin{document}

\maketitle

\begin{reply}
  We thank the referee for the careful review of the text. We discuss below the
changes we have made incorporating your feedback. To make the changes more visible,
in the new draft we have made bold and colored red the parts that are new or have changed.
\end{reply}

@@

\begin{referee}
Major issues :

The authors use a uniform disk model for fitting the extended sources.
However many sources show a complex morphology (for example, not radial symmetric)
They should quantify how the results they are presenting depend on the actual shape of the sources.
\end{referee}

\begin{reply}
In the new draft, Section 2 has new text defining TS.
Section 3 has been renamed (Validation of the TS Distribution) and has been
divided into three sections. The original section is now section 3.1
(now called Point-like Source Simulations Over a Uniform Background).
I will describe section 3.2 in reply to the next question.

There is also a new section 3.3 (Extended Source Simulations Over a 244 Structured Background). 
In this section, we address your question by performing
a dedicated Monte Carlo study of W44 simulated with a ring-type spatial
model. By fitting the source with different spatial models, we assess
this bias and find it to be very small.

This new section requires two new figures (Figure 6 and Figure 7). Also,
in section 2.4 (Comparing Source Size) we add a line of text referring
to this new section. Finally, there is a new sentence in the introduction
of the paper which references the new text.
\end{reply}

@@

\begin{referee}
Most of the new extended sources found in this work lie on the galactic
plane at very low latitudes. Why the simulations used to validate the
analysis (par. 3) do not include the diffuse Galactic background? Can
the structures of this diffuse background affect the analysis results?
\end{referee}

\begin{reply}
In section 3.2 (Point-like Source Simulations Over a Structured Background), we have added new text describing a
follow up Monte Carlo study performed in the presence of the galactic
plane. In addition, this new text required updating Table 1 (Monte Carlo
Spectral Parameters) to include the new simulation parameters. There is also
a new figure (Figure 5) which shows the distribution of $\mathrm{TS}_\mathrm{ext}$
for these
new simulations.
\end{reply}

@@

\begin{referee}
An important result of this work is that some 2FGL sources could be fake due to the residuals induced by modeling the extended sources as point-like. It may be worth having a table that lists the sources that are not longer needed in the modeling of the sky (alternatively this information could be included in table 4). They must also be mentioned in the captions of figures 16, 17, 19, 21 and 22.
\end{referee}

\begin{reply}
We have added a new table (Table 6 in the current draft, named Nearby Residual-induced Sources) which contains
this information. There is also new text in section 9 (Extended Source
Search Method) which refer to this
table. We have also updated the requested captions to list names of 2FGL
sources that were removed.
\end{reply}

@@

\begin{referee}
++++++++++++++++
Report on statistical methodology

Search for Spatially Extended Fermi-LAT Sources Using Two Years of Data

J. Lande et al. (ApJ86244)

a) line 98 states: In the pointlike package, ``The data are binned
spatially, using a HEALPix pixelization, and spectrally (G ́orski et
al. 2005) and the likelihood is maximized over all bins in a region.'' Why
is any binning used? Maximum likelihood estimation (MLE) is best used
on exact photon positions, and binning is not required (as it is with
older chi-square-minimization techniques). Binning is ill-advised as
it results in a loss of information, and sometimes in a bias. As this
assumption is built into pointlike, perhaps it cannot be changed by the
authors for their analysis.
\end{referee}

@@

\begin{reply}
The Fermi LAT analysis package gtlike has an unbinned likelihood mode
which is indeed slightly more precise. But running an unbinned analysis
is very slow because Fermi collects a large number of photons, many of
which do not contribute very much to the detection of a source. gtlike
in binned mode is much faster and is therefore the most common tool
for LAT analysis. Because the LAT collaboration has extensively tested
binned gtlike for spectral analysis, we do not expect it to introduce
any large bias in our spectral analysis result.

But even binned gtlike is prohibitively slow for the extended source
spatial fitting described in this paper. That is why we developed this
new functionality inside of pointlike. 
Even though pointlike performs a fully binned analysis, it used the
healpix hierarchy for spatial bins and scales the spatial bin size with
energy. This ensures that the spatial binning is always smaller than the
inherent spatial resolution of the instrument.
The validations of this method
presented in this paper allow us to be confident in our results.
\end{reply}

@@

\begin{referee}
b) The likelihood ratio test (LRT, eqns 1 \& 9) is a classic technique of
MLE that has been largely replaced by statisticians (since the 1980s) by
closely related measures that take into account the complexity of models
and size of the dataset. The LRT has serious deficiencies: it tends
to accept the more complex model as sample size (i.e., source counts)
increases (or equivalently, it assesses only bias and not variability),
and it is inapplicable when parameters (e.g. source sizes) are near zero
(discussed in the ApJ by Protassov et al. 2002). I also am not confident
that the chi-squared approximation in eqn 8 is correct (please refer to
a statistics book not an old ApJ article, or consult a statistician).

Model selection today is instead guided by `penalized' measures,
and the most widely used is the Bayesian Information Criterion
(BIC,\url{http://en.wikipedia.org/wiki/Bayesian_information_criterion}). It
is very easy to compute, essentially an offset from the LRT scaled to
the log of the counts involved. The issues are discussed in detail in
the volume ``Model Selection and Multi-Model Inference'' (K.P. Burnham
\& D. Anderson, 2nd ed, 2002); see also lectures (especially \#14) at
\url{http://myweb.uiowa.edu/cavaaugh/ms_seminar.html}.

I guess you could argue that your Monte Carlo experiments (Figs 3-4) take
the flux and degrees of freedom into account for the LRT statistic. But
if you want your method to be widely used, it should not rest on a weak
and obsolete statistical foundation. Altogether, I recommend moving to
the BIC. Your approach of Monte Carlo testing of the significance of
LRT values should also work for BIC values.
\end{referee}

\begin{reply}
Concerning the extension significance test, it actually relies not exactly
on Wilk's theorem (which does not hold for the extension test)
but on Chernoff's theorem (see ``On the Distribution of the Likelihood 
Ratio'', \url{http://www.jstor.org/stable/10.2307/2236839}).
we also extensively validated it
against Monte Carlo simulations. Furthermore, the high energy astrophysics
community has a lot of experience with the likelihood ratio test applied to
nested models. So we are very confident that we can apply it to the actual
data.

Concerning the test comparing two point-like sources to one extended
source, we find the desire for a Bayesian test more compelling since the
models are not nested. We considered the BIC test but found that it would
not work well for LAT data. The reason for this is described in the text.
We alternatively considered the AIC test and
found that it is mathematically equivalent to the test we actually performed,
putting the test on firmer ground. We have
added a new paragraph to Section 5 (Testing Against Source Confusion)
that describes this.
\end{reply}

@@

\begin{referee}
Finally, I can suggest that the lead author discuss the issue with experts
at his institution, as Stanford's Dept. of Statistics is universally
recognized as the best in the nation. They might provide a balanced
judgment whether the BIC is substantially better than the LRT in the
situation under study.
\end{referee}

\begin{reply}
Thanks for the suggestion.  In addition to the above changes, we have
fixed a few bibliographical entries.
\end{reply}

### 2012-05-08 extended_catalog/paper/extended_search.tex
update
done w/ draft v10 of ext src paper

 For these reasons, in Section~\ref{analysis_methods_section}
we present a new systematic method for analyzing spatially extended
LAT sources.  
\newtext{
In Section~\ref{monte_carlo_validation}, we demonstrate
that this method can be used to test the statistical significance of the
extension of a LAT source and in Section~\ref{bias_wrong_spatial_model}
we asses the bias that is introduced by assuming an incorrect spatial model.
In Section~\ref{extension_sensitivity},
we calculate the LAT detection threshold to resolve the extension
of a source.}
In Section~\ref{dual_localization_method}, we
study the ability of the LAT to 
distinguish between a single extended source and unresolved closely-spaced point-like sources
In Section~\ref{test_2lac_sources}, we further demonstrate that our
detection method does not misidentify point-like sources as being
extended by testing the extension of active Galactic nuclei (AGN)
believed to be unresolvable.  In Section~\ref{validate_known},
we systematically reanalyze the twelve extended sources included
in the 2FGL catalog and in Section~\ref{systematic_errors_on_extension}
we describe a way to estimate systematic errors on the measured extension of a source.
In Section~\ref{extended_source_search_method}, we
describe a search for new spatially extended LAT sources. Finally,
in Section~\ref{new_ext_srcs_section} we present the detection of the
extension of nine spatially extended sources that were reported in the 2FGL catalog
but treated as point-like in the analysis.  Two of these sources have been previously analyzed in
dedicated publications.

@@

The significance of the extension of a source can be calculated from the
likelihood ratio test of a hypothesis with a spatially extended source and
a hypothesis with a point-like source. The test statistic (TS) for this procedure
is defined as
\begin{equation}
  \tsext=2\log(\likelihood_\text{ext}/\likelihood_\text{ps}) 
\end{equation}
where \likelihood is the Poisson likelihood.
\pointlike calculates \tsext by fitting a source first with a spatially
extended model and then as a point-like source.  The interpretation
of \tsext in terms of a statistical significance is discussed in
Section~\ref{monte_carlo_validation}.

@@

\section{Validation of the Analysis Method}

\subsection{\newtext{Statistical Significance Testing}}

@@

We evaluated the false detection probability for the measurement of source extent
by fitting the extensions of point-like sources.  \cite{mattox_egret}
discuss that the TS distribution for a likelihood ratio test
on the existence of a source at a given position is
\begin{equation}\label{ts_ext_distribution}
  P(\ts)=\onehalf(\chi^2_1(\ts)+\delta(\ts)),
\end{equation}
\newtext{
where $P(\ts)$ is the probability density to get a particular value of TS,
$\chi^2_1$ is the chi-squared distribution with one degree of freedom, and
$\delta$ is the Dirac delta function.
}
The particular form of Equation \ref{ts_ext_distribution} is due to the
null hypothesis (source flux $\Phi=0$) residing on the edge of parameter
space and the model hypothesis adding a single degree of freedom.
It leads to the often quoted result $\sqrt{TS}=\sigma$, where 
$\sigma$ here refers to the significance of the detection. It is plausible
to expect a similar distribution for the TS in the test for
source extension since the same conditions apply (with the source flux
$\Phi$ replaced by the source radius $r$ and $r<0$ being unphysical).
To verify Equation~\ref{ts_ext_distribution}, we evaluated the
empirical distribution function of \tsext computed from simulated sources.

@@

\newtext{
We performed a second simulation to assess any bias that could
be introduces by performing the extension test in the presence of a highly
structured background.
Our simulation setup was the same as above except that the sources were
simulated on top of and analyzed assuming the presence of the standard
Galactic diffuse and isotropic background models used in 2FGL.  In our
simulations, we selected our sources to have a random position on the sky
such that the absolute value of the galactic latitude of the source was
less than 5\degree. This probes the most structured areas of the 
Galactic background.

To limit the number of tests, we selected only one flux
level for each of the four spectral indices and we perform
this test only in the 1 \gev to 100 \gev energy range. 
The fluxes were selected so that $\ts\sim50$. We do not
expect to be able to spatially resolve sources much dimmer than this
and sources much brighter are less likely to be affected by the background.

@@

For each spectrum, we performed $\sim90,000$ simulations.
Figure~\ref{tsext_plane_plot} shows the cumulative density
of \tsext for each spectrum. For small values of \tsext,
there is good agreement between the simulations and
theory.  For the highest values of \tsext, there is possibly
a small discrepancy, but limited statistics makes this hard to quantify.
}

\subsection{Assessing the Bias Due to Incorrect Spatial Models}
\label{bias_wrong_spatial_model}

\newtext{
We performed a Monte Carlo study to quantify the bias that is introduced by
incorrectly modeling the spatial structure of an extended source. 
To assess this systematic, we simulated the spatially extended ring-type SNR W44.
We selected W44 because it is the most significant extended source detected by the LAT 
that has a non-radially symmetric photon-distribution \citep{w44}.

@@

W44 was simulated with the elliptical ring spatial model described in
\cite{w44}. For reference, the ellipse has a semi-major axis of 0\fdg3,
a semi-minor axis of 0\fdg19, a position angle of $147\degree$
measured east of celestial north, and predicts emission only for the
outer 0.75\% of the shape.

We used a simulation setup similar to that described in
Section~\ref{monte_carlo_validation}, but the sources were simulated on
top of and analyzed assuming the presence of the standard Galactic diffuse and
isotropic background model used in 2FGL.  The simulations were
in the 1 \gev to 100 \gev energy range and
W44 was simulated over the two-year time range used in 2FGL. 
In the simulations, we disabled the inherent energy dispersion of photons
by the LAT. The bias caused by ignoring energy dispersion 
will be discussed in an upcoming paper by the LAT collaboration \citep{lat_on_orbit_psf}.
In total, we performed 985 statistically independent simulations.

@@

Figure~\ref{ts_comparison_w44sim}a shows that the simulations of
W44 are very significant ($\ts\sim3500$) when fit as point-like sources.
Figure~\ref{ts_comparison_w44sim}b shows that the simulations 
are highly extended ($\tsext\sim250$)
when fit with a radially-symmetric uniform disk or Gaussian spatial
model.
On average \tsext is
slightly larger when fitting the sources with a uniform disk spatial
model compared to a Gaussian spatial model, consistent with the
fact that the simulated source's true spatial model has a sharp edge.
Figure~\ref{ts_comparison_w44sim}c shows that the sources were fit better
when assuming an elliptical disk spatial model compared to a uniform disk
spatial model ($\ts_\text{elliptical\ disk}-\ts_\text{disk}\sim30$).
Finally, Figure~\ref{ts_comparison_w44sim}d shows that the sources
were fit somewhat better assuming an elliptical ring spatial model
compared to an elliptical disk spatial model ($\ts_\text{elliptical\
ring}-\ts_\text{elliptical\ disk}\sim9$).

Figure~\ref{bias_w44sim}a and Figure~\ref{bias_w44sim}b
show the distributions of best fit fluxes and spectral indices when fitting
the simulated sources.
When fitting the sources with a point-like
spatial model, there is a $\sim$10\% bias in the fit flux and
a $\sim5\%$ bias in the fit index. When fitting the sources assuming any
of the extended spatial models, there is no bias in the best fit spectral parameters
(within the precision of the Monte Carlo study). This is because,
within the finite angular resolution and limited statistics of the LAT, 
all of the extended spatial models can be adjusted to look very similar.

Figure~\ref{bias_w44sim}c plots the best fit 68\% containment radius when
the sources were fit with extended spatial models.  When assuming
the elliptical ring spatial model, there is no significant bias in 
\rsixeight.  Because none of the other spatial models induced a
bias greater than $\sim10\%$, we see that \rsixeight for any of the
spatial models can be used as a reasonable measure of the true 68\%
containment radius for the source.

@@

\tsinc cannot be quantitatively compared to \tsext using a simple
likelihood ratio test to evaluate which model is significantly better
because the models are not nested \citep{statistics_with_care}.
Even though the comparison of \tsext with \tsinc is not a calibrated
test, $\tsext>\tsinc$ indicates that the likelihood for the extended
source hypothesis is higher than for two point-like sources and we only
consider a source to be extended if $\tsext>\tsinc$.

\newtext{
We considered using 
the Bayesian information criterion \citep[BIC,][]{BIC_statistical_test} as
an alternative Bayesian formulation of this test, but it is difficult to apply
to LAT data because it contains a term including the number of data points. For LAT data,
this is not a clear concept because there are many photons included in an analysis that contribute only
 marginally to the analysis of a source.

As an alternative, we considered the Akaike information criterion test \citep[\aic,][]{AIC_statistical_test}.
The \aic is defined as
\begin{equation}
  \aic=2k-2\log\likelihood,
\end{equation}
where $k$ is the number of parameters in the model. 
In this formulation, the best model is considered to be the one that minimizes the \aic.
The first term penalizes models with additional parameters.

@@

For each candidate that was not biased by neighboring point-like
sources or by large-scale residuals in the diffuse emission model, we
improved the model of the region by deciding on a case by case basis which
background point-like sources should be kept.  We kept in our model the
sources that we
believed represented physically distinct sources and we removed 
sources that we believed were included in the 2FGL catalog to compensate
for residuals induced by not modeling the extension of the source.
Soft nearby point-like 2FGL sources that were not significant at higher energies
were frozen to the spectras predicted by 2FGL.
When deciding
which background sources to keep and which to remove, we used as much as possible
multiwavelength information about possibly extended source counterparts
to help guide our choice. For each extended source presented in 
Section~\ref{new_ext_srcs_section}, we describe any modifications from 2FGL
of the background model that were performed.
\newtext{In Table~\ref{fake_2fgl_sources}, we will summarize the sources that were 
included in 2FGL to account for residuals induced by not modeling the extensions of
nearby extended sources.}

@@

Nine extended sources not included in the 2FGL catalog were found by our
extended source search. Two of these have been previously
studied in dedicated publications: RX\,J1713.7$-$3946 and Vela
Jr. \citep{rx_j1713_lat,vela_jr_lat}.
Two of these sources were
found when using photons with energies between 1 \gev and 100 \gev and seven
were found when using photons with energies between 10 \gev and 100 \gev.
For the sources found at energies above 10 \gev, we restrict our
analysis to higher energies because of the issues
of source confusion and diffuse emission modeling described in
Section~\ref{extended_source_search_method}.
The spectral and spatial properties of these nine sources are summarized
in Table~\ref{new_ext_srcs_table} and the results of our investigation of
systematic errors are presented in Table~\ref{alt_diff_model_results}.
Table~\ref{alt_diff_model_results}
also
compares the likelihood assuming the source is spatially
extended to the likelihood assuming 
that the emission originates from two independent point-like
sources. For these new extended sources, $\tsext>\tsinc$ 
so we conclude that 
the \gev emission does not originate from two physically
distinct point-like sources (see Section~\ref{dual_localization_method}).  
Table~\ref{alt_diff_model_results} also includes the
results of the extension fits using variations of the PSF and the Galactic
diffuse model described in Section~\ref{systematic_errors_on_extension}.
There is good agreement between \tsext and the fit size using the standard
analysis, the alternative approach to modeling the diffuse emission, and the alternative PSF.
This suggests that the sources are robust against mis-modeled features in the diffuse
emission model and uncertainties in the PSF.
\newtext{Finally, in Table~\ref{fake_2fgl_sources} we 
summarize the sources included in 2FGL to account for residuals induced
by not modeling the extensions of nearby extended sources.}

@@

\clearpage
\begin{figure}
    \ifcolorfigure
    \plotone{mc_plots/bias_w44sim_color.eps}
    \else
    \plotone{mc_plots/bias_w44sim_bw.eps}
    \fi
    % this plot came from $w44simplots/plot_bias
    \caption{
    \newtext{The distribution of fit parameters
    for the Monte Carlo simulations of W44.
    The plots show the distribution of 
    best fit (a) flux (b) spectral index (c) 68\% containment
    radius (d) eccentricity (e) elliptical orientation angle
    and (f) inner ring fraction.
    }
    }\label{bias_w44sim}
\end{figure}

@@

\tablecomments{
\newtext{
For each new extended sources, we list nearby catalog
sources that were included in 2FGL to account for residuals induced
by not modeling the extension of the nearby extended source.
}
}
\end{deluxetable}

>>>>

 For these reasons, in Section~\ref{analysis_methods_section}
we present a new systematic method for analyzing spatially extended
LAT sources.  
\newtext{
In Section~\ref{monte_carlo_validation}, we demonstrate
that this method can be used to test the statistical significance of the
extension of a LAT source and in Section~\ref{bias_wrong_spatial_model}
we assess the expected level of bias 
introduced by assuming an incorrect spatial model.
In Section~\ref{extension_sensitivity},
we calculate the LAT detection threshold to resolve the extension
of a source.}
In Section~\ref{dual_localization_method}, we
study the ability of the LAT to 
distinguish between a single extended source and unresolved closely-spaced point-like sources
In Section~\ref{test_2lac_sources}, we further demonstrate that our
detection method does not misidentify point-like sources as being
extended by testing the extension of active Galactic nuclei (AGN)
believed to be unresolvable.  In Section~\ref{validate_known},
we systematically reanalyze the twelve extended sources included
in the 2FGL catalog and in Section~\ref{systematic_errors_on_extension}
we describe a way to estimate systematic errors on the measured extension of a source.
In Section~\ref{extended_source_search_method}, we
describe a search for new spatially extended LAT sources. Finally,
in Section~\ref{new_ext_srcs_section} we present the detection of the
extension of nine spatially extended sources that were reported in the 2FGL catalog
but treated as point-like in the analysis.  Two of these sources have been previously analyzed in
dedicated publications.

@@

\newtext{
The significance of the extension of a source can be calculated from the
likelihood-ratio test. The likelihood ratio defines the
test statistic (TS) by comparing the likelihood of a simpler hypothesis to
a more complicated one:
\begin{equation}
  \ts=2\log(\likelihood(H_1)/\likelihood(H_0)),
\end{equation}
where $H_1$ is the more complicated hypothesis and $H_0$ the simpler one.
For the case of the extension test, we compare the likelihood
when assuming the source has either a point-like or spatially extended spatial model:}
\begin{equation}
  \tsext=2\log(\likelihood_\text{ext}/\likelihood_\text{ps}).
\end{equation}
\pointlike calculates \tsext by fitting a source first with a spatially
extended model and then as a point-like source.  The interpretation
of \tsext in terms of a statistical significance is discussed in
Section~\ref{monte_carlo_validation}.

@@

\newtext{
\section{Validation of the \ts Distribution}

\subsection{Point-like Source Simulations Over a Uniform Background}

@@

\newtext{
We tested the theoretical distribution for \tsext
to evaluate the false detection probability for measuring source extension.
To do so, we tested simulated point-like sources for extension.} 
\cite{mattox_egret}
discuss that the \ts distribution for a likelihood-ratio test
on the existence of a source at a given position is
\begin{equation}\label{ts_ext_distribution}
  P(\ts)=\onehalf(\chi^2_1(\ts)+\delta(\ts)),
\end{equation}
\newtext{
where $P(\ts)$ is the probability density to get a particular value of TS,
$\chi^2_1$ is the chi-squared distribution with one degree of freedom, and
$\delta$ is the Dirac delta function.
The particular form of Equation \ref{ts_ext_distribution} is due to the
null hypothesis (source flux $\Phi=0$) residing on the edge of parameter
space and the model hypothesis adding a single degree of freedom (the source extension).}
It leads to the often quoted result $\sqrt{TS}=\sigma$, where 
$\sigma$ here refers to the significance of the detection. It is plausible
to expect a similar distribution for the TS in the test for
source extension since the same conditions apply (with the source flux
$\Phi$ replaced by the source radius $r$ and $r<0$ being unphysical).
To verify Equation~\ref{ts_ext_distribution}, we evaluated the
empirical distribution function of \tsext computed from simulated sources.

@@

\newtext{

\subsection{\newtext{Point-like Source Simulations Over a Structured Background}}

We performed a second simulation to show that the theoretical distribution
for \tsext is still preserved when the point-like sources are present over
a highly-structured diffuse background.
Our simulation setup was the same as above except that the sources were
simulated on top of and analyzed assuming the presence of the standard
Galactic diffuse and isotropic background models used in 2FGL.  In our
simulations, we selected our sources to have a random position on the sky
such that the absolute value of the galactic latitude of the source was
less than 5\degree. This probes the most structured areas of the 
Galactic background.

To limit the number of tests, we selected only one flux
level for each of the four spectral indices and we performed
this test only in the 1 \gev to 100 \gev energy range. 
The fluxes were selected so that $\ts\sim50$. We do not
expect to be able to spatially resolve sources much dimmer than this
and sources much brighter are less likely to be affected by the 
structured background.

@@

For each spectrum, we performed $\sim90,000$ simulations.
Figure~\ref{tsext_plane_plot} shows the cumulative density
of \tsext for each spectrum. For small values of \tsext,
there is good agreement between the simulations and
theory.  For the highest values of \tsext, there is possibly
a small discrepancy, but limited statistics makes this hard to quantify.
Therefore, we are confident we can used \tsext as a robust measure of
statistical significance when testing LAT-detected sources for extension.

\subsection{Extended Source Simulations Over a Structured Background}
\label{bias_wrong_spatial_model}

We also performed a Monte Carlo study to show that incorrectly modeling the
spatial extension of an extended source does not substantially bias
the spectral fit of the source, although it does alter the value of the \ts.
To assess this, we simulated the spatially extended ring-type SNR W44.
We selected W44 because it is the most significant extended source detected by the LAT 
that has a non-radially symmetric photon-distribution \citep{w44}.

@@

W44 was simulated with the elliptical ring spatial model described in
\cite{w44}. For reference, the ellipse has a semi-major axis of 0\fdg3,
a semi-minor axis of 0\fdg19, a position angle of $147\degree$
measured east of celestial north, and the ring's inner radius is 0.75 times
the outer radius.

We used a simulation setup similar to that described in
Section~\ref{monte_carlo_validation}, but the sources were simulated on
top of and analyzed assuming the presence of the standard Galactic diffuse and
isotropic background model used in 2FGL.  The simulations were
in the 1 \gev to 100 \gev energy range and
W44 was simulated over the two-year time range used in 2FGL. 
In the simulations, we disabled the inherent energy dispersion of photons
by the LAT to isolate the effects due to changing the assumed spatial model. 
The bias caused by ignoring energy dispersion 
will be discussed in an upcoming paper by the LAT collaboration \citep{lat_on_orbit_psf}.
In total, we performed 985 statistically independent simulations.

@@

Figure~\ref{ts_comparison_w44sim}a shows that the simulations of W44
are very significant ($\ts\sim3500$) when fit as point-like sources.
Figure~\ref{ts_comparison_w44sim}b shows that the simulations are
highly extended ($\tsext\sim250$) when fit with any of the extended
spatial models.  On average \tsext is somewhat larger when fitting
the sources with more accurate spatial models.  This shows that
assuming an incorrect spatial model will cause the source's
significance to be underestimated.  Figure~\ref{ts_comparison_w44sim}c
shows that the sources were fit better when assuming an elliptical
disk spatial model compared to a uniform disk spatial model
($\ts_\text{elliptical\ disk}-\ts_\text{disk}\sim30$).  Finally,
Figure~\ref{ts_comparison_w44sim}d shows that the sources were
fit somewhat better assuming an elliptical ring spatial model
compared to an elliptical disk spatial model ($\ts_\text{elliptical\
ring}-\ts_\text{elliptical\ disk}\sim9$). This shows that the LAT has
some additional power to resolve substructure in bright extended sources.

Figure~\ref{bias_w44sim}a and Figure~\ref{bias_w44sim}b clearly show that
no significant bias is introduced by modeling the source as extended
but with an inaccurate spatial model, while a point-like source modeling
results in a $\sim10\%$ and $\sim5\%$ bias in the fit flux and index,
respectively.  
Furthermore, Figure~\ref{bias_w44sim}c shows that the \rsixeight estimate of
the extension size is very mildly biased ($\sim10\%$) toward higher values
when inaccurate spatial models are used, and thus represent a reasonable
measurement of the true 68\% containment radius for the source.

@@

\tsinc cannot be quantitatively compared to \tsext using a simple
likelihood-ratio test to evaluate which model is significantly better
because the models are not nested \citep{statistics_with_care}.
Even though the comparison of \tsext with \tsinc is not a calibrated
test, $\tsext>\tsinc$ indicates that the likelihood for the extended
source hypothesis is higher than for two point-like sources and we only
consider a source to be extended if $\tsext>\tsinc$.

\newtext{
We considered using 
the Bayesian information criterion \citep[BIC,][]{BIC_statistical_test} as
an alternative Bayesian formulation for this test, but it is difficult to apply
to LAT data because it contains a term including the number of data points. For LAT data,
this is not a clear concept because there are many photons included in an analysis that contribute only
 marginally to the analysis of a given source.

As an alternative, we considered the Akaike information criterion test \citep[\aic,][]{AIC_statistical_test}.
The \aic is defined as
\begin{equation}
  \aic=2k-2\log\likelihood,
\end{equation}
where $k$ is the number of parameters in the model. 
In this formulation, the best hypothesis is considered to be the one that minimizes the \aic.
The first term penalizes models with additional parameters.

@@

For each candidate that was not biased by neighboring point-like
sources or by large-scale residuals in the diffuse emission model, we
improved the model of the region by deciding on a case by case basis which
background point-like sources should be kept.  We kept in our model the
sources that we
believed represented physically distinct sources and we removed 
sources that we believed were included in the 2FGL catalog to compensate
for residuals induced by not modeling the extension of the source.
Soft nearby point-like 2FGL sources that were not significant at higher energies
were frozen to the spectras predicted by 2FGL.
When deciding
which background sources to keep and which to remove, we used as much as possible
multiwavelength information about possibly extended source counterparts
to help guide our choice. For each extended source presented in 
Section~\ref{new_ext_srcs_section}, we describe any modifications from 2FGL
of the background model that were performed.
\newtext{In Table~\ref{fake_2fgl_sources}, we summarize the sources that were 
included in 2FGL to account for residuals induced by not modeling the extensions of
nearby extended sources.}

@@

Nine extended sources not included in the 2FGL catalog were found by our
extended source search. Two of these have been previously
studied in dedicated publications: RX\,J1713.7$-$3946 and Vela
Jr. \citep{rx_j1713_lat,vela_jr_lat}.
Two of these sources were
found when using photons with energies between 1 \gev and 100 \gev and seven
were found when using photons with energies between 10 \gev and 100 \gev.
For the sources found at energies above 10 \gev, we restrict our
analysis to higher energies because of the issues
of source confusion and diffuse emission modeling described in
Section~\ref{extended_source_search_method}.
The spectral and spatial properties of these nine sources are summarized
in Table~\ref{new_ext_srcs_table} and the results of our investigation of
systematic errors are presented in Table~\ref{alt_diff_model_results}.
Table~\ref{alt_diff_model_results}
also
compares the likelihood assuming the source is spatially
extended to the likelihood assuming 
that the emission originates from two independent point-like
sources. For these new extended sources, $\tsext>\tsinc$ 
so we conclude that 
the \gev emission does not originate from two physically
distinct point-like sources (see Section~\ref{dual_localization_method}).  
Table~\ref{alt_diff_model_results} also includes the
results of the extension fits using variations of the PSF and the Galactic
diffuse model described in Section~\ref{systematic_errors_on_extension}.
There is good agreement between \tsext and the fit size using the standard
analysis, the alternative approach to modeling the diffuse emission, and the alternative PSF.
This suggests that the sources are robust against mis-modeled features in the diffuse
emission model and uncertainties in the PSF.

@@

\clearpage
\begin{figure}
    \ifcolorfigure
    \plotone{mc_plots/bias_w44sim_color.eps}
    \else
    \plotone{mc_plots/bias_w44sim_bw.eps}
    \fi
    % this plot came from $w44simplots/plot_bias
    \caption{
    \newtext{The distribution of fit parameters
    for the Monte Carlo simulations of W44.
    The plots show the distribution of 
    best fit (a) flux (b) spectral index (c) 68\% containment
    radius (d) eccentricity (e) elliptical orientation angle
    and (f) inner ring fraction. The dashed vertical lines represent
    the simulated values of the parameters.
    }
    }\label{bias_w44sim}
\end{figure}

@@

\tablecomments{
\newtext{
For each new extended source, we list nearby catalog
sources that were included in 2FGL to account for residuals induced
by not modeling the extension of the extended source.
}
}
\end{deluxetable}
